---
title: "IC Trends"
output: html_document
---

```{r}
# Import the data 
IC = read.csv("IC by job 1.2015 - 4.2017 no dup - 4.28 - CSV.csv")


```


*Data Preparation*

```{r}

# Regression Tree
library(rpart)
library(rpart.plot)
library(class)

# Interaction 
library(effects)
library(ggplot2)

# Split the data into training and validation groups
set.seed(1234)
IC_rand = IC[sample(nrow(IC)),]
cutoff = round(0.7*nrow(IC))
IC_train = IC_rand[1:cutoff,]
IC_test = IC_rand[-(1:cutoff),]

```

##################################################################
##################################################################
##################################################################


*Data Exploration*

# Summarizing the data

```{r}
head(IC)
```

```{r}
# Calculate Mean and SD by Move Type


df <- data.frame(MoveType=IC$MoveType, CF=IC$CF, Mileage=IC$Mileage, Revenue=IC$Revenue, LaborCost=IC$LaborCost, Profit=IC$Profit, 
                                        LaborCostPerCF=IC$CostPerCF, LaborToRev=IC$LaborToRev)

aggregate(.~MoveType, df, function(x) c(mean=mean(x)))
```

```{R}
aggregate(.~MoveType, df, function(x) c(sd=sd(x)))

```



```{r}
ggplot(IC, aes(x=MoveType, y=Profit, fill=MoveType)) + geom_boxplot() +
  ggtitle("Profit by Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32"),
        text = element_text(size=8.5, face = "bold")) +
  guides(fill=FALSE)

```


```{r}
ggplot(IC, aes(x=MoveType, y=Profit.Margin, fill=MoveType)) + geom_boxplot() +
  ggtitle("Profit Margin by Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32"),
        text = element_text(size=8.5, face = "bold")) +
  guides(fill=FALSE)

```



```{R}
ggplot(IC, aes(x=MoveType, y=LaborToRev, fill=MoveType)) + geom_boxplot() +
  ggtitle("LabToRev by Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32"),
        text = element_text(size=9, face = "bold")) +
  guides(fill=FALSE)
```


```{r}

plot(IC$Profit ~ IC$JobDate)

```


##################################################################
##################################################################
##################################################################

*Regression Tree*

```{r}
# Build a regression tree model 
tree = rpart(Profit ~ MoveType + Mileage + CF, data=IC)

summary(tree)

# Validate the regression tree using 'Complexity Parameter'
printcp(tree)

plotcp(tree)

ptree <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]), "CP"])

# Make a fancy decision tree
library(rattle)

fancyRpartPlot(ptree, uniform=TRUE, tweak =1.5, cex=0.5)

```




Evaluate the Decision Tree

```{r}
treepred <- predict(tree, newdata=IC_test)

head(treepred)

library(caret)
#confusionMatrix(treepred, IC_test$Profit)

```


Try *party* package (unbiased for mixed types of variables)

```{r}

library(party)
tree2 = ctree(Profit ~ MoveType + Mileage + CF, data=IC_train)

plot(tree2, type="simple")


```



##################################################################
##################################################################
##################################################################


*Regression Model*

Check Assumptions (to see if linear model would fit)

First of all, check with the pairwise scatterplot. There seems to be *no curvilinear relationship* between IC and any potential predictors. Also, we can find out that there might be some interactions between 

```{r}
pairs(~ Profit + CF + MoveType + Mileage, data=IC)


```


Full Model Selection 

- Not too much differences between these two models (residual)
- Decided to use the second model 

```{r}

IC$CF <- as.numeric(IC$CF)

IC$Mileage <- as.numeric(IC$Mileage)

# Full Mocel
lm.full <- lm(Profit ~ CF*Mileage*MoveType, data=IC)
drop1(lm.full, test="F")

```

Interaction term *Mileage:MoveType* is significant at the 10% significance level, so I chose the following linear model: 

__Profit ~ CF + Mileage + MoveType + Mileage:MoveType__


```{r}

# compute lower and upper whiskers
ylim1 = boxplot.stats(IC$Mileage)$stats[c(1, 5)]

ggplot(IC, aes(x=MoveType, y=Mileage, fill=MoveType)) + geom_boxplot() +
  ggtitle("Mileage by Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32"),
        text = element_text(size=8.5, face = "bold")) + coord_cartesian(ylim = ylim1*1.05) +
  guides(fill=FALSE)


```






```{r}
ggplot(IC, aes(x=MoveType, y=CF, fill=MoveType)) + geom_boxplot() +
  ggtitle("Cubic Feet by Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32"),
        text = element_text(size=8.5, face = "bold")) +
  guides(fill=FALSE)
```



```{r}

# Check Assumption 

plot(density(resid(lm.full))) #density plot 
par(mfrow=c(2,2))
plot(lm.full)


# Transformation 
  ## Box-Cox Transformation 
  ## Log Transformation 

```

*Visualizing for Interaction*

```{r}

# save the coefficients into a vector
cs <- coef(lm.full)

# Profit vs. Mileage vs. MoveType
ggplot(IC, aes(x=Mileage, y=Profit, color=MoveType)) + geom_point() + 
  geom_abline(intercept = cs[1], slope = cs[2]) + 
  geom_smooth(method="lm") +
  ggtitle("Profit by Mileage and Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32")) +
  xlim(0,1000)
```


```{r}
# Profit vs. CF vs. MoveType

ggplot(IC, aes(x=CF, y=Profit, color=MoveType)) + geom_point() + 
  geom_abline(intercept = cs[1], slope = cs[2]) + 
  geom_smooth(method="lm") +
  ggtitle("Profit by Cubic Feet and Move Type \n(IC; July 2015 - April 2017)") +
  theme(plot.title=element_text(face="bold", size=14, hjust=0.5, color="gray32"))
```



```{r}
lm.test <- lm(Profit.Margin ~ CF*Mileage*MoveType, data=IC)
drop1(lm.test, test="F")
plot(allEffects(lm.test))

```


```{r}
summary(lm.full)
```


__71.5%__ variability of the *Profit* data is explained by *CF*, *Mileage*, and *MoveType* variables. 

*Linear Model Interpretation*


*Linear Model Prediction*

```{r}

pred=predict(lm.full, 
             newdata=data.frame(CF=402, Mileage=26.2, MoveType="Local Moving"),
             se.fit=TRUE)
z.perc=qnorm(0.90)
LB=pred$fit - z.perc*pred$se.fit
UB=pred$fit + z.perc*pred$se.fit
cbind(LB, UB)

```



*Coefficient Interpretation*

```{r}
# confint.default(lm.full)

coef(lm.full)

```


